---
title: 'Can money buy happiness?'
author: "Jocelyn Radke, Ana Espeleta, Cat Terres"
output: 
  html_document:
    keep_md: TRUE
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    code_download: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

```{r}
#Loading the packages
library(dplyr)
library(readr)
library(ggplot2)
library(gridExtra)
library(tidyverse)
library(lubridate) 
library(ggthemes)
library(plotly)
library(ggdark)
library(maps)
library(ggmap)
library(leaflet)
library(RColorBrewer)
```

## Introduction and Background

Can money buy happiness? Philosophers, university students, and lay people alike have grappled with this question for ages. The knee-jerk reaction is to emphatically reject a link between life satisfaction and financial abundance because everyone knows both miserable wealthy people and others who appear incredibly joyful despite severe financial difficulties. As one rapper (The Notorious B.I.G.) insightfully observed that, “Mo Money Mo Problems,” which illustrates that affluence prompts challenges of its own. Some researchers have turned to empirical testing to seek an answer to this question. One study by [Jachimowicz](https://www.hbs.edu/ris/Publication%20Files/FinancialScarcity_SPPS_accepted_November1_2021_06f87dd3-4474-4c92-88d1-a5c6b63a2eaa.pdf) and colleagues found that financial instability correlates with higher levels of daily distress. Although money alone doesn’t imply happiness, money can eliminate daily hardships that lead to life dissatisfaction. A 2010 [Princeton](https://spia.princeton.edu/sites/default/files/content/docs/news/Happiness_Money_Summary.pdf) study measured that a $75,000 annual income was the “sweet spot” for maximizing emotional well-being. Incomes below this benchmark were correlated with lower levels of happiness and higher levels of stress. These research studies and our own analysis seek to make a connection between standards of living (a quantitative measure) and quality of life (a qualitative measure). A standard of living measures the material well-being of people in a given geographical population which is reflected in factors such as comfort, access to opportunities, goods, healthcare services, and safety. Quality of life, on the other hand, is a more subjective term that reflects the level of happiness of a population, which certainly varies person to person.

> “Wealth consists not in having great possessions, but in having few wants.” – Epictetus, stoic philosopher

Economic performance is a common determinate of standard of living within a country. The reasoning goes that a country that functions well economically is able to provide stable income and access to jobs, increase the availability of goods and services, and afford a social safety net. The most common measure of standard of living is gross domestic product (GDP), which is a measure of the total volume of goods and services produced within in country in a given year. We paired GDP data with population data to calculate GDP per capita (thereafter, GDPPC) so as to control for differences in population between countries. We believe that GDPPC provides a helpful, though imperfect measure of the standard of living within a country. GDPPC operates with the assumption that every individual within a country gets an equal slice of the GDP pie, which is unrealistic because of widespread income inequality, however, can provide insight as long as we are mindful of its limitations. In what follows, we seek to challenge the assumption that the best countries are those that are most economically successful though looking at measures of quality of living in addition to standard of living. 

## Data Collection

The data set used is titled [Healthy Lifestyle Cities Report 2021](https://www.kaggle.com/datasets/prasertk/healthy-lifestyle-cities-report-2021) and was imported from Kaggle.com in a csv file. It contains a total of 44 cities, which were selected and ranked by its creator (a U.K.-based company), using 10 different metrics. For our study, we decided to focus on two of the variables provided, happiness index by country and annual hours worked by city. We supplemented this data set with a data set we created that included GDP data, population data, and weekly hours worked. We gathered our GDP data from the [International Monetary Fund's (IMF) World Economic Outlook database](https://www.imf.org/en/Publications/WEO/weo-database/2018/April). The average work week by country data came from an article in the [World Population Review](https://worldpopulationreview.com/country-rankings/average-work-week-by-country) that sourced the data to the Penn World Table 2019 data. 


**Happiness Index**

The “Happiness Levels” variable was taken directly from a specific pillar of the [Happy Planet Index](https://happyplanetindex.org/countries/), which was generated using survey data from the [World Happiness Report](https://worldhappiness.report/ed/2021/). The collection of survey data is described on the Happy Planet Index site and consists of asking people to rate their current life on a scale from 0-10. Their perception of the best possible life corresponds to a 10 on the scale, and their perception of the worst possible life to a 0. The final scores are computed using weighted averages by population and consider the scores from 1-2 years prior to increase the sample size and give more precise estimates. The poll gathers around 1,000 responses annually for each country.


## Analysis

```{r}
#Loading the data
healthy_data <- read_csv("healthy_lifestyle_city_2021.csv", 
    col_types = cols(Rank = col_number(), 
        `Sunshine hours(City)` = col_number(), 
        `Cost of a bottle of water(City)` = col_number(), 
        `Obesity levels(Country)` = col_number(), 
        `Life expectancy(years) (Country)` = col_number(), 
        `Pollution(Index score) (City)` = col_number(), 
        `Annual avg. hours worked` = col_number(), 
        `Happiness levels(Country)` = col_number(), 
        `Outdoor activities(City)` = col_number(),
        `Number of take out places(City)` = col_number(), 
        `Cost of a monthly gym membership(City)` = col_number()))

supplemental_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vQL_6se7aBLQSP92RgP5DkvQ_quTyzsTxskf9ZFpvE_5ifMS5whQ1Lr0JRNMtd1ZGAaVXlEEMECPMHr/pub?output=csv", 
    col_types = cols(GDPM = col_number(), 
        GDP = col_number(), pop = col_number(), 
        GDPPC = col_number()))

#Joining data and converting pounds to dollars
healthy_data_2 <- healthy_data %>% 
  left_join(supplemental_data,
            by = c("City" = "city")) %>% 
  mutate(water_bottle_cost = `Cost of a bottle of water(City)`*1.3,
         gym_cost = `Cost of a monthly gym membership(City)`*1.3)
```

The following map shows the cities that were chosen in the analysis.

```{r}
labels <- sprintf(
  "<strong>%s</strong><br/> %s",
  healthy_data_2$City, healthy_data_2$country) %>%
  lapply(htmltools::HTML)

leaflet(data = healthy_data_2) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    radius = 4,
    color = "#D4A200",
    stroke = FALSE, 
    fillOpacity = 0.7,
    label = ~labels
  )
```

> “Wealth is like sea-water: The more we drink the thirstier we become; and the same is true of fame.” -Schopenhauer


```{r}
GDPP_hours <- healthy_data_2 %>% 
ggplot( aes(x = GDPPC, y = `Hours Worked`, color = `City`)) +
  geom_point() +
  labs(title = "Gross Domestic Product per Capita by Country and Average Weekly Hours Worked",
       caption = "Data from: ADD LATER",
       x = "GDP per capita (USD)",
       y = "Hours Worked (Weekly)",
       color = "") +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "lightblue",
                                colour = "gray",
                                size = 0.5,
                                ),
        plot.title = element_text(size = 0.05))


ggplotly(GDPP_hours,
         width = 800,
         height = 430,
         tooltip = c("text", 
                     "x",
                     "City"))
```

This initial graph explores the relationship between GDP per capita and the average number of weekly hours worked in each city. It allows us to understand what these two factors tell us about the overall well-being of each city later in the study.
The plot presents a predominantly inverse relationship between GDP per capita and the number of weekly hours worked. A cluster of cities is in the range of $0-$20K GDP per capita, and except for one city (Johannesburg), all can be classified as Latin American or Asian. We can infer that this relationship may describe different production systems and processes in each country. A lower GDP and higher number of hours worked likely points to a lack of access to technological resources or poor organizational management which makes it difficult to increase the standard of living. Countries with higher levels of technology can produce the same output as a country with lower levels of technology with much fewer inputs of labor or machinery hours.
The second observable cluster on this plot is within the bounds of $40K-$60K GDP per capita, with a range of 12.2 between the lowest and highest number of weekly working hours. However, most points in the 30-35 weekly working hour range correspond to European cities, which follow the trend of higher GDP per capita and lower working hours. In contrast to our previous analyses, the increased GDP may represent more efficiency of labor, which can sustain the economy with fewer working hours and thus, promote a higher standard of living. 


```{r}
options(scipen=999)

ggplot(healthy_data_2, aes(y = fct_reorder(country, GDPPC), x = GDPPC, color = `Happiness levels(Country)`)) +
  geom_point() +
  labs(title = "Gross Domestic Product per Capita by Country and Happiness Level",
       caption = "Data from: ADD LATER",
       x = "GDP per capita (USD)",
       y = "",
       color = "Happiness Level 
       (1-10 scale)") +
  theme(plot.title = element_text(hjust = 1)) +
  theme_classic() +
  scale_color_gradient(high = "#FFD25A", low = "black")
```

We chose to investigate if there is a correlation between GDP per capita and happiness levels by country. We speculated that countries with a higher GDPPC might be less happy reflecting longer work hours required to maintain such a high level of production. However, the GDPPC versus hours worked graph indicated a loose negative relationship between GDPPC and working hours, which seems to imply that people in countries with higher levels of national production actually work less hours. This second graph shows that higher levels of GDP appear to correlate to higher life satisfaction. This implies that higher standards of living and higher material wealth may translate to more happiness.


> “I think that this is ultimately why we want money: It's to create a world that you prefer." -Hank Green 



```{r, fig.width=5,fig.alt= "This graph plots various cities as their own color coded dot with a caption when scrolled over. Variables happiness and hours worked are on the x and y axes, respectively. The captions attached to each point contain happiness level, average weekly hours worked, and the name of the city. Following the graph from left to right, a loose trend line appears amongst the points." }

healthy_data_dot <- healthy_data_2 [-c(23),] #Getting rid of Egypt because missing average work week data

Healthy <- healthy_data_dot %>% 
  select(`City`, "Happiness Level" = `Happiness levels(Country)`, "Average Weekly Hours Worked" = `Hours Worked`) %>% 
  mutate(`Avg. Weekly Hours Worked` = factor(`Average Weekly Hours Worked`)) %>% 
  
 ggplot(aes(x = `Happiness Level`, 
             y = `Avg. Weekly Hours Worked`, 
             color = `City`)) +
  geom_point() +
  scale_y_discrete(breaks = c(26.6, 30.6, 33.2, 38.8, 42.1))+
  
  labs(title = "Is there a correlation between weekly work hours and happiness levels?",
       x = "Happiness Level",
       y = "",
       caption = "Graph By Cat Terres; data sourced from Kaggle, 
       'Healthy Lifestyle Cities Report 2021'",
       color = "") +
  
  theme_foundation() +
  theme(legend.title=element_blank())+
  theme(legend.background = element_blank()) +
  theme(legend.position = "none")
  

ggplotly(Healthy,
         width = 800,
         tooltip = c("text",
                     "y",
                     "x",
                     "City")) %>% 
  layout(title = list(text = paste0('Is there a correlation between weekly work hours and happiness levels?',
                                    '<br>',
                                    '<sup>',
                                    'Graph by Cat Terres, data sourced from Kaggle and World Population Review',
                                    '</sup>')))
```

This graph simply presents reported happiness level against the average weekly hours worked. Our marked interest in working hours out of all the factors in our initial dataset stems from what we witness each and every day in our lives, communities, and the nation at large. It is easy to assume that those who work less have more time for leisure activities and maintenance of their well-being, and in turn we speculated that fewer weekly work hours would correlate to a greater happiness level. However, this graph does not present a strong enough trend to support our hypothesis. There is a loose trend, and this can help to give direction to our investigation despite not proving anything for a fact. What we DO learn from this graph is the complexity of happiness as a variable. Happiness is not only contingent on working hours, but factors such as higher average wages, government investment in citizens and their communities, and access to community resources. We recognize that there are cultural, political, and societal nuances that influence whether people work more or less and that different places have different attitudes surrounding work and money. 

> “It is not the man who has too little, but the man who craves more, that is poor.” – Seneca



>Concluding Statements: Can money buy happiness? Our key findings from this project are ______________. We sought to tell a story about the connection between economic prosperity and happiness but with the added factor of working hours. Money is inextricable from the work needed to obtain it, and to base our argument purely on GDP or how much money is possessed completely ignores a key factor: how much are people working to obtain said money. Thus, we tied the two together, factoring in that having more money likely ______ and working fewer hours means that people will _____________. This presents a paradox of sorts, more work generally means more money, and more money means more happiness, but less work also generates more happiness. So where is a harmonious balance found? We see that ________. Our research cannot possibly cover all factors such as access to community programs, optimal living conditions, or types of government and the length to which they go to provide resources. All of these ancillary factors play a role in happiness and the ability to work fewer hours, but we posit that looking purely at money and work can still produce illuminating discoveries. 